% ============================================================
%  quiz.tex — Inside a Smartphone: The Tiny World in Your Hand
%  End-of-book quiz — 6 parts + answer key, matching Book 1 structure.
% ============================================================

\namedchapter{Quiz Time!}
\addcontentsline{toc}{chapter}{Quiz Time!}

\chapteropening{%
  You have explored every corner of a smartphone — from the processor
  to the GPS satellites far above! Now test what you have learned.
}

% ============================================================
\section*{Part 1: True or False?}
% ============================================================

\begin{tcolorbox}[enhanced,colback=SkyBlue!10,colframe=TechPurple,arc=6pt,boxrule=1pt,
  left=10pt,right=10pt,top=8pt,bottom=8pt]

Write \textbf{T} (true) or \textbf{F} (false) next to each statement.

\begin{enumerate}[label=\textbf{\arabic*.}]
  \item A smartphone is just a phone — it cannot do anything a computer can do. \hfill \rule{1.5cm}{0.4pt}
  \item The processor carries out instructions and controls the whole phone. \hfill \rule{1.5cm}{0.4pt}
  \item RAM keeps your photos saved even when the phone is switched off. \hfill \rule{1.5cm}{0.4pt}
  \item Touchscreens detect the heat from your finger. \hfill \rule{1.5cm}{0.4pt}
  \item A camera sensor converts light into digital information. \hfill \rule{1.5cm}{0.4pt}
  \item Batteries store chemical energy and convert it into electricity. \hfill \rule{1.5cm}{0.4pt}
  \item The gyroscope tells your phone which direction is north. \hfill \rule{1.5cm}{0.4pt}
  \item GPS works by receiving signals from satellites orbiting Earth. \hfill \rule{1.5cm}{0.4pt}
  \item You need at least three satellites to find your exact position. \hfill \rule{1.5cm}{0.4pt}
  \item Using the same password for every app is a good idea. \hfill \rule{1.5cm}{0.4pt}
\end{enumerate}
\end{tcolorbox}

\vspace{12pt}

% ============================================================
\section*{Part 2: Match the Character to Their Job}
% ============================================================

\begin{tcolorbox}[enhanced,colback=WarmOrange!8,colframe=WarmOrange,arc=6pt,boxrule=1pt,
  left=10pt,right=10pt,top=8pt,bottom=8pt]

Draw a line from each character to what they do inside the phone:

\begin{center}
\begin{tabular}{p{0.32\linewidth}p{0.05\linewidth}p{0.52\linewidth}}
\textbf{Character} & & \textbf{Job} \\[8pt]
1. Dot   & & A. Runs your favourite apps and games \\[6pt]
2. Chip  & & B. A returning guide — a packet of information \\[6pt]
3. Memo  & & C. Captures photos and videos \\[6pt]
4. Tappy & & D. The brain that follows all instructions \\[6pt]
5. Lens  & & E. Stores photos and apps permanently \\[6pt]
6. Zap   & & F. Detects the electrical charge in your fingertip \\[6pt]
7. Sensa & & G. Powers the phone with stored energy \\[6pt]
8. Appy  & & H. Feels every tilt, twist, and direction change \\[6pt]
9. Orbi  & & I. Orbits Earth and helps the phone find its location \\[6pt]
\end{tabular}
\end{center}
\end{tcolorbox}

\vspace{12pt}

% ============================================================
\section*{Part 3: Fill in the Blanks}
% ============================================================

\begin{tcolorbox}[enhanced,colback=LeafGreen!8,colframe=LeafGreen,arc=6pt,boxrule=1pt,
  left=10pt,right=10pt,top=8pt,bottom=8pt]

Use the word bank to complete each sentence.

\begin{center}
\begin{tcolorbox}[colback=SunYellow!20,colframe=SunYellow!70!black,arc=4pt,boxrule=0.6pt,width=0.9\linewidth]
\centering\bfseries
RAM \quad processor \quad accelerometer \quad satellite \quad pixels \quad operating system
\end{tcolorbox}
\end{center}

\begin{enumerate}[label=\textbf{\arabic*.}]
  \item The \rule{3cm}{0.4pt} is the brain of the phone that carries out billions of instructions every second.
  \item \rule{3cm}{0.4pt} is the working memory that holds information the phone is actively using.
  \item A photo is made up of millions of tiny dots of colour called \rule{3cm}{0.4pt}.
  \item GPS works by receiving signals from at least three \rule{3cm}{0.4pt}s orbiting Earth.
  \item The \rule{3cm}{0.4pt} measures how fast your phone is speeding up or slowing down.
  \item The \rule{3cm}{0.4pt} is the main programme (like Android or iOS) that manages the whole phone.
\end{enumerate}
\end{tcolorbox}

\vspace{12pt}

% ============================================================
\section*{Part 4: Put It in Order}
% ============================================================

\begin{tcolorbox}[enhanced,colback=SoftPurple!8,colframe=SoftPurple,arc=6pt,boxrule=1pt,
  left=10pt,right=10pt,top=8pt,bottom=8pt]

Number these steps 1 to 6 to show what happens when you take a photo and save it:

\begin{center}
\begin{tabular}{cl}
\rule{1cm}{0.4pt} & The processor compresses the image file. \\[6pt]
\rule{1cm}{0.4pt} & You tap the shutter button on the screen. \\[6pt]
\rule{1cm}{0.4pt} & The image is saved to internal storage. \\[6pt]
\rule{1cm}{0.4pt} & Light passes through the lens onto the camera sensor. \\[6pt]
\rule{1cm}{0.4pt} & The camera sensor converts light into digital information. \\[6pt]
\rule{1cm}{0.4pt} & The touchscreen detects your finger and sends the command. \\[6pt]
\end{tabular}
\end{center}
\end{tcolorbox}

\vspace{12pt}

% ============================================================
\section*{Part 5: Smartphone Word Challenge}
% ============================================================

\begin{tcolorbox}[enhanced,colback=CoralRed!6,colframe=CoralRed,arc=6pt,boxrule=1pt,
  left=10pt,right=10pt,top=8pt,bottom=8pt,
  title={\faPuzzlePiece\quad Smartphone Word Challenge},fonttitle=\bfseries\large]

\textbf{Across clues:}
\begin{enumerate}[label=\textbf{\arabic*.}]
  \item The brain of the phone that follows instructions (9 letters)
  \item A network of orbiting machines used to find your location (3 letters)
  \item The working memory that holds what the phone is doing right now (3 letters)
  \item A programme that runs on your phone to do a specific job (3 letters)
  \item A tiny dot of colour in a photo or on a screen (5 letters)
\end{enumerate}

\vspace{8pt}
\textbf{Down clues:}
\begin{enumerate}[label=\textbf{\arabic*.}]
  \item The permanent memory where photos and apps are saved (7 letters)
  \item A sensor that measures how the phone is tilting or rotating (9 letters)
  \item The part of the camera that focuses light (4 letters)
  \item A device that orbits Earth and sends GPS signals (9 letters)
  \item A device that stores energy and powers the phone (7 letters)
\end{enumerate}
\end{tcolorbox}

\vspace{12pt}

% ============================================================
\section*{Part 6: Think About It}
% ============================================================

\begin{tcolorbox}[enhanced,colback=SunYellow!10,colframe=SunYellow!70!black,arc=6pt,boxrule=1pt,
  left=10pt,right=10pt,top=8pt,bottom=8pt]

Answer these questions in your own words:

\begin{enumerate}[label=\textbf{\arabic*.}]
  \item Why does a phone slow down when you open lots of apps at the same time?
  \item GPS works even when there is no Wi-Fi. How is that possible?
  \item Why do smartphone cameras take better photos in daylight than at night?
  \item You download a new game and your phone says ``Storage full.''  What could you do?
  \item What are two ways to keep your smartphone and your personal information safe?
\end{enumerate}
\end{tcolorbox}

\vspace{16pt}

% ============================================================
\section*{Answer Key}
% ============================================================

\begin{tcolorbox}[enhanced,colback=LightGray,colframe=TechPurple!50,arc=6pt,boxrule=1pt,
  left=10pt,right=10pt,top=8pt,bottom=8pt,
  title={Answers (no peeking until you try!)},fonttitle=\bfseries\color{DarkText}]

\textbf{Part 1:}\; 1.\,F \quad 2.\,T \quad 3.\,F \quad 4.\,F \quad 5.\,T \quad 6.\,T \quad 7.\,F \quad 8.\,T \quad 9.\,T \quad 10.\,F

\vspace{6pt}
\textbf{Part 2:}\; 1-B \quad 2-D \quad 3-E \quad 4-F \quad 5-C \quad 6-G \quad 7-H \quad 8-A \quad 9-I

\vspace{6pt}
\textbf{Part 3:}\; 1.\,processor \quad 2.\,RAM \quad 3.\,pixels \quad 4.\,satellite \quad 5.\,accelerometer \quad 6.\,operating system

\vspace{6pt}
\textbf{Part 4:}\; 3, 1, 6, 2, 4, 5

\vspace{6pt}
\textbf{Part 5 Across:}\; 1.\,PROCESSOR \quad 2.\,GPS \quad 3.\,RAM \quad 4.\,APP \quad 5.\,PIXEL

\textbf{Part 5 Down:}\; 1.\,STORAGE \quad 2.\,GYROSCOPE \quad 3.\,LENS \quad 4.\,SATELLITE \quad 5.\,BATTERY

\end{tcolorbox}
