% ============================================================
%  quiz.tex — The Secret Life of the Internet
%  End-of-book quiz and review activities.
% ============================================================

\namedchapter{Quiz Time!}
\addcontentsline{toc}{chapter}{Quiz Time!}

\chapteropening{%
  You have explored the whole internet — from Wi-Fi waves to undersea cables!
  Now test your knowledge with these fun challenges.
}

% ============================================================
\section*{Part 1: True or False?}
% ============================================================

\begin{tcolorbox}[enhanced,colback=SkyBlue!10,colframe=InternetBlue,arc=6pt,boxrule=1pt,
  left=10pt,right=10pt,top=8pt,bottom=8pt]

Write \textbf{T} (true) or \textbf{F} (false) next to each statement.

\begin{enumerate}[label=\textbf{\arabic*.}]
  \item The internet is one single giant computer. \hfill \rule{1.5cm}{0.4pt}
  \item Wi-Fi uses radio waves to connect devices. \hfill \rule{1.5cm}{0.4pt}
  \item A URL is the home address of a website. \hfill \rule{1.5cm}{0.4pt}
  \item Data travels between continents mostly by satellite. \hfill \rule{1.5cm}{0.4pt}
  \item A router directs traffic between your devices and the internet. \hfill \rule{1.5cm}{0.4pt}
  \item Packets always travel the same route to reach their destination. \hfill \rule{1.5cm}{0.4pt}
  \item The cloud is actually computers in buildings on the ground. \hfill \rule{1.5cm}{0.4pt}
  \item HTML is the language used to build web pages. \hfill \rule{1.5cm}{0.4pt}
  \item A search engine visits every page on the internet to build its index. \hfill \rule{1.5cm}{0.4pt}
  \item You should use the same password for all your accounts. \hfill \rule{1.5cm}{0.4pt}
\end{enumerate}
\end{tcolorbox}

\vspace{12pt}

% ============================================================
\section*{Part 2: Match the Character to Their Job}
% ============================================================

\begin{tcolorbox}[enhanced,colback=WarmOrange!8,colframe=WarmOrange,arc=6pt,boxrule=1pt,
  left=10pt,right=10pt,top=8pt,bottom=8pt]

Draw a line from each character to what they do:

\begin{center}
\begin{tabular}{p{0.35\linewidth}p{0.05\linewidth}p{0.5\linewidth}}
\textbf{Character} & & \textbf{Job} \\[8pt]
1. Dot & & A. Carries signals through the air \\[6pt]
2. Bee Browser & & B. Stores web pages and sends them to you \\[6pt]
3. Willa Wi-Fi & & C. A tiny packet of information \\[6pt]
4. Rory Router & & D. Carries data under the ocean \\[6pt]
5. Sunny Server & & E. Takes you to any website \\[6pt]
6. Captain Cable & & F. Helps you find information \\[6pt]
7. Scout Search & & G. Directs traffic in and out of your home \\[6pt]
8. Cloud Pal & & H. Stores your files on faraway computers \\[6pt]
\end{tabular}
\end{center}
\end{tcolorbox}

\vspace{12pt}

% ============================================================
\section*{Part 3: Fill in the Blanks}
% ============================================================

\begin{tcolorbox}[enhanced,colback=LeafGreen!8,colframe=LeafGreen,arc=6pt,boxrule=1pt,
  left=10pt,right=10pt,top=8pt,bottom=8pt]

Use the word bank to complete each sentence.

\begin{center}
\begin{tcolorbox}[colback=SunYellow!20,colframe=SunYellow!70!black,arc=4pt,boxrule=0.6pt,width=0.85\linewidth]
\centering\bfseries
packets \quad\quad DNS \quad\quad optical fibres \quad\quad router \quad\quad crawlers \quad\quad TCP
\end{tcolorbox}
\end{center}

\begin{enumerate}[label=\textbf{\arabic*.}]
  \item When you type a web address, your browser asks the \rule{3cm}{0.4pt} to find the server's number address.
  \item Information on the internet is split into tiny pieces called \rule{3cm}{0.4pt}.
  \item The \rule{3cm}{0.4pt} in your home connects all your devices and sends requests to the internet.
  \item Undersea cables carry data as pulses of light through \rule{3cm}{0.4pt}.
  \item Search engines use programmes called \rule{3cm}{0.4pt} to visit and read web pages.
  \item \rule{3cm}{0.4pt} is the rule that makes sure all packets arrive and none are lost.
\end{enumerate}
\end{tcolorbox}

\vspace{12pt}

% ============================================================
\section*{Part 4: Put It in Order}
% ============================================================

\begin{tcolorbox}[enhanced,colback=SoftPurple!8,colframe=SoftPurple,arc=6pt,boxrule=1pt,
  left=10pt,right=10pt,top=8pt,bottom=8pt]

Number these steps 1 to 6 to show what happens when you visit a website:

\begin{center}
\begin{tabular}{cl}
\rule{1cm}{0.4pt} & Your browser displays the page on your screen. \\[6pt]
\rule{1cm}{0.4pt} & You type a web address and press Enter. \\[6pt]
\rule{1cm}{0.4pt} & The server sends back the page in pieces (packets). \\[6pt]
\rule{1cm}{0.4pt} & Your browser asks DNS for the server's IP address. \\[6pt]
\rule{1cm}{0.4pt} & Your browser contacts the server. \\[6pt]
\rule{1cm}{0.4pt} & DNS replies with the number address. \\[6pt]
\end{tabular}
\end{center}
\end{tcolorbox}

\vspace{12pt}

% ============================================================
\section*{Part 5: Vocabulary Challenge}
% ============================================================

\begin{tcolorbox}[enhanced,colback=CoralRed!6,colframe=CoralRed,arc=6pt,boxrule=1pt,
  left=10pt,right=10pt,top=8pt,bottom=8pt,
  title={\faPuzzlePiece\quad Internet Word Challenge},fonttitle=\bfseries\large]

\textbf{Across clues:}
\begin{enumerate}[label=\textbf{\arabic*.}]
  \item A programme that lets you visit websites (7 letters)
  \item The company that brings the internet to your home (3 letters)
  \item A collection of pages stored on a server (7 letters)
  \item Invisible waves that carry Wi-Fi signals (5, 5 letters)
  \item The internet's ``phone book'' that turns names into numbers (3 letters)
\end{enumerate}

\vspace{8pt}
\textbf{Down clues:}
\begin{enumerate}[label=\textbf{\arabic*.}]
  \item A tiny chunk of data travelling through the internet (6 letters)
  \item A powerful computer that stores and sends information (6 letters)
  \item A thin glass thread that carries light signals (7, 5 letters)
  \item A device that directs traffic between your home and the internet (6 letters)
  \item Servers accessed via the internet to store your files remotely (5 letters)
\end{enumerate}
\end{tcolorbox}

\vspace{12pt}

% ============================================================
\section*{Part 6: Think About It}
% ============================================================

\begin{tcolorbox}[enhanced,colback=SunYellow!10,colframe=SunYellow!70!black,arc=6pt,boxrule=1pt,
  left=10pt,right=10pt,top=8pt,bottom=8pt]

Answer these questions in your own words:

\begin{enumerate}[label=\textbf{\arabic*.}]
  \item Why is it faster to watch a video from a server nearby than from one on the other side of the world?
  \item What would happen if all the undersea cables were cut at the same time?
  \item Why is the internet sometimes called a ``network of networks''?
  \item If the cloud is not really in the sky, why do you think people call it that?
  \item What are two things you can do to stay safe online?
\end{enumerate}
\end{tcolorbox}

\vspace{16pt}

% ============================================================
\section*{Answer Key}
% ============================================================

\begin{tcolorbox}[enhanced,colback=LightGray,colframe=InternetBlue!50,arc=6pt,boxrule=1pt,
  left=10pt,right=10pt,top=8pt,bottom=8pt,
  title={Answers (no peeking until you try!)},fonttitle=\bfseries\color{DarkText}]

\textbf{Part 1:} 1.\,F \quad 2.\,T \quad 3.\,T \quad 4.\,F \quad 5.\,T \quad 6.\,F \quad 7.\,T \quad 8.\,T \quad 9.\,F \quad 10.\,F

\vspace{6pt}
\textbf{Part 2:} 1-C \quad 2-E \quad 3-A \quad 4-G \quad 5-B \quad 6-D \quad 7-F \quad 8-H

\vspace{6pt}
\textbf{Part 3:} 1.\,DNS \quad 2.\,packets \quad 3.\,router \quad 4.\,optical fibres \quad 5.\,crawlers \quad 6.\,TCP

\vspace{6pt}
\textbf{Part 4:} 6, 1, 5, 2, 4, 3

\vspace{6pt}
\textbf{Part 5 Across:} 1.\,BROWSER \quad 2.\,ISP \quad 3.\,WEBSITE \quad 4.\,RADIO WAVES \quad 5.\,DNS

\textbf{Part 5 Down:} 1.\,PACKET \quad 2.\,SERVER \quad 3.\,OPTICAL FIBRE \quad 4.\,ROUTER \quad 5.\,CLOUD
\end{tcolorbox}
