% ============================================================
%  glossary.tex — Robot World: Meet the Machines That Help Us
% ============================================================

\namedchapter{Glossary}
\addcontentsline{toc}{chapter}{Glossary}

\begin{description}[labelwidth=4.2cm, leftmargin=4.4cm]

  \item[\glossterm{Accelerometer}] A sensor that measures changes in speed or direction.

  \item[\glossterm{Actuator}] A part that moves or does work \textemdash{} motors, grippers, propellers.

  \item[\glossterm{Assembly line}] A production system where products move past workstations, each adding a part.

  \item[\glossterm{Autonomous}] Able to act without direct human control, making decisions independently.

  \item[\glossterm{Algorithm}] A step-by-step method for solving a problem.

  \item[\glossterm{Artificial intelligence (AI)}] Computer systems that perform tasks normally requiring human thought.


  \item[\glossterm{AUV}] Autonomous Underwater Vehicle \textemdash{} navigates on its own underwater.

  \item[\glossterm{Battery}] A portable device that stores electrical energy for a robot to use.

  \item[\glossterm{Autonomy}] The ability to act without human control.

  \item[\glossterm{Bio-inspired}] Designed by copying ideas from nature.

  \item[\glossterm{Boundary wire}] A buried wire that tells a garden robot where to stop mowing.

  \item[\glossterm{Chassis}] The main frame or body of a robot.

  \item[\glossterm{Charging dock}] A station where a robot returns to refill its battery.

  \item[\glossterm{Cobot}] A collaborative robot designed to work safely alongside humans.

  \item[\glossterm{Companion robot}] A robot designed for social interaction and emotional support.

  \item[\glossterm{Condition}] A yes/no check that changes what a programme does next.

  \item[\glossterm{Controller}] The computer board inside a robot that runs the programme.

  \item[\glossterm{Conveyor belt}] A moving belt that carries items through a factory.


  \item[\glossterm{Data privacy}] Controlling who can see your personal information.

  \item[\glossterm{Degrees of freedom}] The number of independent ways a robot can move.

  \item[\glossterm{Drone}] An unmanned flying robot.

  \item[\glossterm{Emotion detection}] Using AI to guess how someone is feeling.

  \item[\glossterm{Ethics}] The study of what is right and wrong.

  \item[\glossterm{Energy}] The power needed to make motors spin, sensors work, and controllers think.

  \item[\glossterm{Exoskeleton}] A wearable robotic frame that supports or strengthens the body.

  \item[\glossterm{Face recognition}] Technology that identifies a person from their facial features.

  \item[\glossterm{Gear}] A toothed wheel that transfers force between shafts.

  \item[\glossterm{Gripper}] A robot hand designed to grab objects.


  \item[\glossterm{Infrared}] Invisible light used by sensors to detect heat or movement.

  \item[\glossterm{Gyroscope}] A sensor that measures rotation and tilt.

  \item[\glossterm{Infrared}] Invisible light used by sensors to detect heat or movement.

  \item[\glossterm{Humanoid}] A robot shaped like a human body.

  \item[\glossterm{Joint}] A connection between two parts that allows bending or rotating.

  \item[\glossterm{Lidar}] Light Detection and Ranging \textemdash{} uses lasers to map distances.

  \item[\glossterm{Loop}] An instruction that repeats until a condition is met.

  \item[\glossterm{Machine learning}] A type of AI where computers improve through experience.

  \item[\glossterm{Microcontroller}] A small, low-cost computer designed to control a single machine or device.

  \item[\glossterm{Micro-bot}] A very small robot, often smaller than a fingernail.

  \item[\glossterm{Motor}] A device that converts electricity into spinning motion.

  \item[\glossterm{Nano-bot}] A theoretical robot at the scale of molecules.

  \item[\glossterm{Programme}] A set of instructions written in a programming language.

  \item[\glossterm{Programming language}] A language humans use to write instructions for computers (e.g.\\ Python, Scratch).

  \item[\glossterm{Propeller}] A spinning blade that pushes air (or water) to move a drone or underwater robot.

  \item[\glossterm{Radiation}] Invisible energy in space that is harmful to living things.

  \item[\glossterm{Robot}] A machine that can sense, think, and act on its own.

  \item[\glossterm{ROS}] Robot Operating System \textemdash{} a popular framework for robot software.

  \item[\glossterm{ROV}] Remotely Operated Vehicle \textemdash{} controlled by a human via cable.

  \item[\glossterm{Rover}] A wheeled robot designed to travel across a planet's surface.

  \item[\glossterm{Safety cage}] A fence around a factory robot to protect human workers.

  \item[\glossterm{Sensor}] A device that measures something in the environment.

  \item[\glossterm{Sensor fusion}] Combining data from multiple sensors for a better picture.

  \item[\glossterm{Signal delay}] The time a radio message takes to travel a long distance.

  \item[\glossterm{Soft robot}] A robot made of flexible materials that can bend and stretch.

  \item[\glossterm{Solar panel}] A panel that converts sunlight into electricity.

  \item[\glossterm{Stop button}] An emergency control that immediately halts a robot's movement.

  \item[\glossterm{Surgical robot}] A robot system that helps surgeons perform operations.

  \item[\glossterm{Swarm robotics}] Many simple robots cooperating to accomplish complex tasks.

  \item[\glossterm{Teleoperation}] Controlling a robot from a distance using radio signals or a cable.

  \item[\glossterm{Telesurgery}] Surgery performed remotely using a robot.

  \item[\glossterm{3D printing}] Building objects layer by layer from a digital model.

  \item[\glossterm{Ultrasonic}] Using sound waves above human hearing to detect objects.

  \item[\glossterm{Ultraviolet (UV)}] Invisible light used to kill bacteria and viruses.

  \item[\glossterm{Welding}] Joining two metal pieces together using intense heat.

\end{description}
